# GEOVIZ

<img src="img/rgeoviz.svg" style="width:150px; display:block; border:none;" alt="Logo">

`geoviz` is an R package for thematic mapping. As its name suggests, it's an implementation of the [geoviz](https://github.com/riatelab/geoviz) javascript library, itself based on the [d3.js](https://d3js.org/) ecosystem ported by Mike Bostock. Like the original javascript library, the package can be used to create a wide range of interactive, zoomable vector maps, taking advantage of d3's many features: proprtional symbols, pictograms, typologies, choropleth maps, Spikes, tiles, Droing cartograms, etc. It can also be used to create pretty static vectorial maps in svg format, suitable for editorial cartography.  

# Principles

- The `geoviz` package is an R implementation of the `geoviz` JavaScript library via a htmlwidget. Its development follows the evolution of the library. The parameter names are the same. You can therefore refer to the geoviz documentation [here](https://riatelab.github.io/geoviz/). Therefore, d3geoviz is not intended to compete with other mapping packages in R, such as [mapsf](https://cran.r-project.org/web/packages/mapsf/index.html) or [tmap](https://cran.r-project.org/web/packages/tmap/index.html).

- Since it's based on d3.js, the philosophy behind this package is completely different. Map parameters use svg attributes rather than the usual R parameters. Thus, `strokeWidth` is used rather than `lwd`, `fill` rather than `col`, `stroke` rather than `border`, etc.

- `geoviz` is not designed to handle voluminous datasets. It is suitable for light, generalized basemaps. 

- `geoviz` is designed to work with geographic data in wgs84 (not projected). Geometries are then projected on the fly using the `create()` function. Unlike other R packages based on `sf`, the projections used come from the d3.js ecosystem (d3-geo, d3-geo-projection & d3-geo-polygon).

- Maps generated by `geoviz` are zoomable.  Two types of zoom are available. The classic type (pan and zoom) and the “versor” type for creating interactive globes.

- Maps created with `geoviz` are interactive. It is therefore possible to create tooltips to access information contained in geographic objects.

- Many different types of maps are available. The types can be combined with each other and are highly customizable. 


# Here an example


```{r, message=FALSE}
library("sf")
world <- st_read(system.file("gpkg/world.gpkg", package = "geoviz"), quiet = TRUE)
```

```{r, message=FALSE}
viz_create() |>
  viz_outline() |>
  viz_graticule() |>
  viz_path(data = world) |>
  viz_header(text = "Map Title") |>
  viz_footer(text = "Source, author, year, etc.") |>
  viz_render()
  ```
