---
title: "create and render"
author: "Nicolas Lambert"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{create and render}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, echo=FALSE, message=FALSE}
library(geoviz)
library(sf)
world <- st_read(system.file("gpkg/world.gpkg", package = "geoviz"), quiet = TRUE)
```

# Create

Every map starts with the `viz_create` function. It initializes the map and defines a number of general parameters, such as the size (`width`), the margins (`margins`), the projection (`projection`), or the background color (`background`). This is also where you determine whether the map is zoomable or not, using the `zoomable` parameter.

From a technical perspective, the function produces a list containing the map parameters and an initially empty list of layers to be displayed.

```{r}
map <- viz_create(projection = "EqualEarth", margin = c(20,0,0,0), background = "#55bee0")
str(map)
```

At this stage, the function does not draw anything.

# Add layers

Once the map has been initialized with the `viz_create` function, the geoviz package provides many functions to display layers. Each function has its own specific parameters and is designed to display a particular type of layer. For example, you can use `viz_path` to add gemeotries, `viz_graticule` to add graticules, or `viz_header` to add a title, etc. 

So you can write this

```{r}
map <- viz_create(projection = "EqualEarth")
map <- viz_outline(map)
map <- viz_graticule(map, stroke = "white", step =20)
map <- viz_path(map, data = world)
map <- viz_header(map, text = "Hello World")
str(map)
```

Or that (recomended)

```{r}
map <- viz_create(projection = "EqualEarth") |>
 viz_outline() |>
 viz_graticule(stroke = "white", step =20) |>
 viz_path(data = world) |>
 viz_header(text = "Hello World")
str(map)
```

At this stage, you still do not draw anything. You have created a configuration file containing the data and parameters needed to build the map.

# Render

This configuration file will be rendered using the `viz_render()` function.

```{r}
viz_render(map)
```

Thus, creating a complete map with geoviz could be written as follows.

```{r, eval = FALSE}
map <- viz_create(projection = "EqualEarth") |>
 viz_outline() |>
 viz_graticule(stroke = "white", step =20) |>
 viz_path(data = world) |>
 viz_header(text = "Hello World") |>
 viz_render()
```

From a technical perspective, the viz_render function returns an HTML block (+ JavaScript for interactivity) containing a map in SVG format.

# Save maps

The maps created with the geoviz package thus generate a web page. You can save this web page with the `htmlwidgets` package:

```{r, eval = FALSE}
library("htmlwidgets")
```

```{r, eval = FALSE}
viz_create(zoomable = T) |>
  viz_path(data = world, tip = T) |>
  viz_render() |>
  saveWidget("mymap.html", selfcontained = TRUE)
```

You can also directly save the map in SVG format to edit it with a drawing software like Inkscape. For this, you just need to use the `viz_save` command instead of `viz_render`.

```{r, eval = FALSE}
viz_create() |>
  viz_path(data = world) |>
  viz_save("mymap.svg")
```

By doing this, the map is returned and a window opens, prompting you to download it. Thus, the geoviz package can be a way to create draft maps before refining them in a CAD environment.

