---
title: "effects"
author: "Nicolas Lambert"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{effects}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, echo=FALSE, message=FALSE}
library(geoviz)
library(sf)
world <- st_read(system.file("gpkg/world.gpkg", package = "geoviz"), quiet = TRUE)
aus <- world[world$ISO3 == "AUS",]
```

The `geoviz` package makes it possible to create publication-quality maps. Since these maps are built in SVG format, SVG filters can be defined. This works in two steps. First, you define an effect by giving it an identifier. Then, you apply it to the layer of your choice. 

# blur

The `viz_blur` function is used to define a blur effect The `id` parameter defines the filter identifier. The `stdDeviation` parameter determines the blur range.

```{r}
# A simple map
viz_create() |>
  # 1 - create filter
  viz_blur(id = "my_blur_effect", stdDeviation =  5) |>
  # 2 - add path and apply filter
  viz_path(datum = aus, fill =  "#38896F", filter = "url(#my_blur_effect)") |>
  viz_render()
```

# shadow

The `viz_shadow` function can be used to draw a drop shadow

```{r}
# A simple map
viz_create(margin = 5) |>
  viz_shadow(id = "my_shadow_effect", stdDeviation =  2.5, dx = 5, dy = 5) |>
  viz_path(datum = aus, fill =  "#38896F", filter =  "url(#my_shadow_effect)") |>
  viz_render()
```

# radialGradient

With `viz_radialGradient`, you can create a radial gradient. In addition to the filter id, the function takes several parameters as input. `color1` and `color2` define the gradient colors. The parameters `offset1`, `offset2`, `fx` and `fy` define how the gradation is performed. Then, you have to apply it on the `fill` attribute.

```{r}
# A simple map
viz_create(margin = 5) |>
  viz_radialGradient(id = "my_radial_gradient", color1 = "#7db054", color2 = "#38896f
", offset1 = 50, offset2 = 100, fx = 50, fy = 50) |>
  viz_path(datum = aus, fill = "url(#my_radial_gradient)") |>
  viz_render()
```

# clipPath

The <clipPath> SVG element defines a clipping path, to be used by the clip-path property. A clipping path restricts the region to which paint can be applied. In geoviz, you can add a clipPath element with the `viz_clipPath()` function.

```{r}
# A simple map
viz_create() |>
  viz_clipPath(id = "my_clip_path", datum =  aus )|>
  viz_tile(url = "worldimagery", clipPath = "url(#my_clip_path)") |>
  viz_render()
```

# pattern

The `viz_pattern` function allows you to create patterns and apply them to any geometry. It supports multiple textures: `"lines"`, `"crosshatch"`, `"dots"`, `"waves"`, `"triangles"`, and `"zigzag"`.

```{r}
# A simple map
viz_create() |>
  viz_pattern(data = aus, pattern = "cross", stroke = "#38896F", strokeOpacity = 2, angle = 45) |>
  viz_render()
```

# sketch

The `viz_sketch` function generates a hand-drawn (sketchy) SVG representation of a spatial dataframe. It applies SVG filters (*feTurbulence* and *feDisplacementMap*) to create a pencil-like effect. 

```{r}
# A simple map
viz_create() |>
  viz_sketch(data = aus, stroke = "#38896F") |>
  viz_render()
```
