---
title: "Hello geoviz"
author: "Nicolas Lambert"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Hello geoviz}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r}
#| include: false
knitr::opts_chunk$set(out.width = "100%")
```

The `geoviz` package is a (partial) R interface to the [geoviz](https://github.com/riatelab/geoviz) JavaScript library. It allows users to create vector-based, interactive, and zoomable thematic maps. The package is therefore built on the web ecosystem, relying on SVG and the [D3.js](https://d3js.org/) library. As a result, some argument names may be slightly confusing for R users. In addition, the package is designed for relatively lightweight basemaps and is not intended for handling large datasets.

Here an example of a map designed with the package:

```{r, message=FALSE, echo=FALSE}
library(geoviz)
library(sf)
world <- st_read(system.file("gpkg/world.gpkg", package = "geoviz"), quiet = TRUE)
africa <- world[world$region == "Africa", ]
viz_create(zoomable = TRUE, projection = "mercator", domain = africa, margin = 20) |>
  viz_shadow(id = "ombre") |>
  viz_tile(url = "worldphysical") |>
  viz_header(text = "Hello Geoviz", background_fill = "#38896F", fill = "white", fontSize = 30) |>
  viz_graticule(stroke = "black", strokeWidth = 1.5, strokeOpacity = 0.3, strokeDasharray = 5, transform = "translate(1,1)") |>
  viz_graticule(stroke = "white", strokeWidth = 1.5, strokeOpacity = 0.5, strokeDasharray = 5) |>
  viz_choro(data = africa, var = "gdppc", tip = "$NAMEen\n\nGDP per capita: $gdppc\nPopulation:$pop", fillOpacity = 0.8, stroke = "none", leg_pos = c(40, 390), leg_title = "GDP per Capita", leg_frame = TRUE, leg_frame_stroke = "none", filter = "url(#ombre)") |>
  viz_prop(data = africa, var = "pop", stroke = "white", strokeWidth = 1.3, fill = "none", leg_pos = c(40, 577), leg_title = "Population", leg_frame = TRUE, leg_frame_stroke = "none") |>
  viz_scalebar() |>
  viz_north(scale = 1.5, pos = c(663, 155), fill = "#38896F") |>
  viz_render()
```

# Installation

To install the package, just write:

```{r, eval = FALSE}
install.packages("remotes")
remotes::install_github("riatelab/geoviz_R")
```

then

```{r}
library(geoviz)
```

# Syntax

There are several steps involved in creating a map with `geoviz`. 

**1** - First, create the map container using the **`create()`** function. This is where you define the projection, margins, background color, etc. - in short, all the general parameters of the map.

**2** The next step is to progressively add layers. A set of dedicated functions is available for this purpose. For instance, `viz_path` adds a spatial dataframe, `viz_graticule` draws latitude and longitude lines, `viz_header` inserts a title, and `viz_footer` adds a source note. All available functions are documented in the Reference section.


**3** - Then, the **`render()`** function displays the map

The various functions can be chained together with a pipe (**`|>`**).

Now, let's build a world map.

```{r, message=FALSE}
# Data Loading
library(sf)

world <- st_read(
  system.file("gpkg/world.gpkg", package = "geoviz"),
  quiet = TRUE
)

cities <- st_read(
  system.file("gpkg/cities.gpkg", package = "geoviz"),
  quiet = TRUE
)
```

*⚠️ The geometries must be in latitude and longitude coordinates, not projected. Use WGS84.*

```{r}
viz_create() |>
  viz_outline() |>
  viz_graticule() |>
  viz_path(data = world) |>
  viz_path(data = cities) |>
  viz_header(text = "Hello World") |>
  viz_render()
```

# Styling

As you can see, the name of the parameters don't look like classical R parameters. Since it's based on d3.js, map parameters use svg attributes rather than the usual R parameters. Thus, `strokeWidth` is used rather than `lwd`, `fill` rather than `col`, `stroke` rather than `border`, etc. 

All layers have default parameters, but you can modify them. Here is a list of parameters you can use on most layers:

- `fill`: Sets the fill color of the element. Can be a fixed color (e.g., "steelblue") or a function mapping a data variable to colors.
- `stroke`: Sets the color of the element’s outline (border). Example: "#000" for black.
- `strokeWidth`: Width of the outline in pixels. Controls how thick the border appears.
- `opacity`: Overall opacity of the element, from 0 (fully transparent) to 1 (fully opaque).
- `fillOpacity`: Opacity of the fill only, separate from the stroke.
- `strokeOpacity`: Opacity of the stroke only.
- `strokeDasharray`: Defines a dash pattern for the stroke. Example: "5,5" creates dashed lines of 5px segments with 5px gaps.
- `strokeLinecap`: Defines the shape of the ends of lines. Values: "butt", "round", "square".
- `strokeLinejoin`: Defines the shape of joints between connected lines. Values: "miter", "round", "bevel".

*⚠️ Be careful: pay attention to uppercase and lowercase letters.*

Let’s take the previous map and customize it to our liking.

```{r}
viz_create() |>
  viz_outline(fill = "#38896F", fillOpacity = 0.3) |>
  viz_graticule(stroke = "white", step = 20) |>
  viz_path(
    data = world, fill = "white",
    stroke = "none", fillOpacity = 0.3
  ) |>
  viz_path(
    data = cities, pointRadius = 2,
    stroke = "none", fill = "#38896F"
  ) |>
  viz_header(
    text = "Hello World", fill = "white",
    background_fill = "#38896F"
  ) |>
  viz_render()
```

# Positioning of elements

To position symbols or text on the map, we most often use the coordinates from a spatial dataframe. However, it is also possible to place specific elements at precise coordinates. For this, two coordinate systems coexist in geoviz:

- **Geographical coordinates**

By specifying `coords = "geo"`, you can position an element at precise geographic coordinates. The object is then linked to the map and stays in the correct location according to the zoom level.

```{r}
# A simple map
mumbai <- c(72.88, 19.07)
viz_create(zoomable = T) |>
  viz_path(data = world, fill = "#CCC", stroke = "none", fillOpacity = 0.3) |>
  viz_circle(pos = mumbai, coords = "geo", fill = "#38896F", r = 10) |>
  viz_render()
```

- **SVG coordinates**

By specifying `coords = "svg"`, you can also position an element statically directly on the map plane. This is very useful for decorative elements, such as the legend or text labels.

**⚠️ <ins>The origin of the coordinate system used to express coordinates in an SVG document is located at the top left.</ins>**

```{r}
# A simple map
viz_create(zoomable = T) |>
  viz_path(data = world, fill = "#CCC", stroke = "none", fillOpacity = 0.3) |>
  viz_circle(pos = c(300, 100), coords = "svg", fill = "#38896F", r = 10) |>
  viz_render()
```

# Symbology

Specialized functions are available for representing statistical data, whether qualitative or quantitative. For example:

- `viz_choro`: for relative quantitative data
- `viz_prop`: for absolute quantitative data
- `viz_typo`: for qualitative data

These functions take two main parameters: `data` for the base map and `var` for the variable to be mapped. Additional parameters allow for fine customization of both the map and its legend.

Here is a simple example of a proportional symbol map

```{r}
viz_create() |>
  viz_path(datum = world, fill = "#CCC") |>
  viz_prop(data = world, var = "pop", fill = "#38896F") |>
  viz_render()
```

A choropleth map

```{r}
viz_create() |>
  viz_choro(data = world, var = "gdppc", fill = "#38896F") |>
  viz_render()
```

And a typology

```{r}
viz_create() |>
  viz_typo(data = world, var = "region") |>
  viz_render()
```

For more details, see the “symbology” page.




