---
title: "layers"
author: "Nicolas Lambert"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{layers}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, echo=FALSE, message=FALSE}
library(geoviz)
library(sf)
world <- st_read(system.file("gpkg/world.gpkg", package = "geoviz"), quiet = TRUE)
```

The geoviz library offers numerous functions that allow you to display specific layers and overlay them to create the map of your choice. We list them below.

# Base Map and Structure

## Path

The `viz_path` function allows you to display any basemap (points, lines, and polygons). You can customize the display using SVG parameters such as `fill`, `stroke`, `strokeWidth`, `opacity`, `fillOpacity`, and [many others](https://riatelab.github.io/geoviz/global.html#path).

```{r}
viz_create(projection = "EqualEarth") |>
  viz_path(data = world, fill = "#38896F") |>
  viz_render()
```

## Outline

The `viz_outline` function allows to create a layer with Earth outline in the projection. You can customize the display using SVG parameters such as `fill`, `stroke`, `strokeWidth`, `opacity`, `fillOpacity`, and [many others](https://riatelab.github.io/geoviz/global.html#outline).

```{r}
viz_create(projection = "EqualEarth") |>
  viz_outline(fill = "#38896F") |>
  viz_path(data = world, fill = "white", fillOpacity = 0.2, stroke = "none") |>
  viz_render()
```

## Graticule

The `viz_graticule` function allows function allows to create a layer with lat/long lines. You can customize the display using SVG parameters such as `step`, `stroke`, `strokeWidth`, `strokeDasharray`, and [many others](https://riatelab.github.io/geoviz/global.html#graticule).

```{r}
viz_create(projection = "EqualEarth") |>
  viz_path(data = world, fill = "#38896F", fillOpacity = 0.2, stroke = "none") |>
  viz_graticule(step = 20, stroke = "#38896F", strokeDasharray = c(5,10)) |>
  viz_render()
```

## Earth

The `viz_earth` function displays PNG files representing the Earth's surface. The data comes from Natural Earth. Credit as follows: Natural Earth. Free vector and raster map data @ naturalearthdata.com. Unlike tiles, you can choose whichever projection you want. the `url` parameter allows to change the image to display. You can choose one from this list: `"GRAY_50M_SR"`, `"GRAY_50M_SR_OB"`, `"GRAY_50M_SR_W"`, `"HYP_50M_SR"`, `"HYP_50M_SR_W"`, `"MSR_50M"`, `"NE1_50M_SR_W"`, `"NE2_50M_SR"`, `"NE2_50M_SR_W"`, `"OB_50M"`, `"PRIMSA_SR_50M"`, `"SR_50M"`. See Other parameters [here](https://riatelab.github.io/geoviz/global.html#earth).  

```{r}
viz_create(projection = "Polar") |>
  viz_earth(url = "HYP_50M_SR_W") |>
  viz_render()
```

## Tile

The `viz_tile` function allows function allows to add zoomable raster tiles. When you use this mark, the Mercator projection is automaticaly used. With the parmeter `url`, you can choose the style of the tiles. Try `"openstreetmap"`, `"opentopomap"`, `"worldterrain"`, `"worldimagery"`, `"worldStreet"`, `"worldphysical"`, `"shadedrelief"`, `"stamenterrain"`, `"cartodbvoyager"`, `"stamentoner"` , `"stamentonerbackground"` , `"stamentonerlite"`, `"stamenwatercolor"`, `"hillshade"`, `"worldocean"`, `"natgeo"` or `"worldterrain"`. See Other parameters [here](https://riatelab.github.io/geoviz/global.html#tile).

```{r}
viz_create() |>
  viz_tile(url = "worldimagery") |>
  viz_render()
```

# Map Decorations and Annotations

## Labels

The `viz_text` function allows to add a text on the map. It allow also to create a layer with labels from a spatial dataframe. See all parameters [here](https://riatelab.github.io/geoviz/global.html#text).

```{r}
africa <- world[world$region == "Africa",]
viz_create(projection = "Mercator", domain = africa) |>
  viz_path(datum = africa, fill = "#CCC", fillOpacity = 0.5) |>
  viz_text(data = africa, text = "name", fill = "#38896F", stroke = "white", strokeWidth = 4) |>
  viz_render()
```

## Header

The `viz_header` function allows function allows to add a title above the map. You can customize the display using parameters such as `text`, `fill`, `fill_background`, `textAnchor`, `dominantBaseline` and [many others](https://riatelab.github.io/geoviz/global.html#header).

```{r}
viz_create(projection = "EqualEarth") |>
  viz_outline(fill = "#CCCCCC") |>
  viz_path(data = world, fill = "white", fillOpacity = 0.2, stroke = "none") |>
  viz_header(text = "Hello World", fill = "#38896F") |>
  viz_render()
```

## Footer

The `viz_footer` function allows function allows to add a text below the map. You can customize the display using parameters such as `text`, `fill`, `fill_background`, `textAnchor`, `dominantBaseline` and [many others](https://riatelab.github.io/geoviz/global.html#footer).

```{r}
viz_create(projection = "EqualEarth") |>
  viz_outline(fill = "#CCCCCC") |>
  viz_path(data = world, fill = "white", fillOpacity = 0.2, stroke = "none") |>
  viz_footer(text = "Author, source, date...", fill = "#38896F") |>
  viz_render()
```

# North

The `viz_north` function adds a north arrow. By default, the arrow indicates true north at the location where it is placed. Depending on the map projection, the arrow may therefore not point in the same direction everywhere on the map. The `pos` parameter allows you to position the arrow using the SVG document coordinates (0,0 is at the top-left corner). See all parameters [here](https://riatelab.github.io/geoviz/global.html#north).

```{r}
africa <- world[world$region == "Africa",]
viz_create(projection = "Mercator", domain = africa) |>
  viz_path(datum = africa, fill = "#CCC", fillOpacity = 0.5) |>
  viz_north(pos = c(50,50), fill = "#38896F") |>  
  viz_render()
```

## Scalebar

The `viz_scalebar` function allows add a scalebar. The `pos` parameter allows you to position the scalebar. See all parameters [here](https://riatelab.github.io/geoviz/global.html#scalebar).

```{r}
viz_create(projection = "Mercator", domain = africa) |>
  viz_path(datum = africa, fill = "#CCC", fillOpacity = 0.5) |>
  viz_scalebar() |>  
  viz_render()
```


