---
title: "Layout"
author: "Nicolas Lambert"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Layout}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r}
#| include: false
knitr::opts_chunk$set(out.width = "100%")
```

```{r, echo=TRUE, message=FALSE}
library(geoviz)
library(sf)

world <- st_read(
  system.file("gpkg/world.gpkg", package = "geoviz"),
  quiet = TRUE
)
```

The geoviz library offers numerous functions that allow you to display specific layers and overlay them to create the map of your choice. This gives you plenty of options for laying out your map.

# Base Map and Structure

## Outline

The `viz_outline` function allows to create a layer with Earth outline in the projection. You can customize the display using SVG parameters such as `fill`, `stroke`, `strokeWidth`, `opacity`, `fillOpacity` and many others.

```{r}
viz_create(projection = "EqualEarth") |>
viz_outline(fill = "#38896F") |>
viz_path(data = world, fill = "white",
         fillOpacity = 0.2, stroke = "none") |>
viz_render()
```

## Graticule

The `viz_graticule` function allows function allows to create a layer with lat/long lines. You can customize the display using SVG parameters such as `step`, `stroke`, `strokeWidth`, `strokeDasharray`, and many others.

```{r}
viz_create(projection = "EqualEarth") |>
viz_path(data = world, fill = "#38896F",
         fillOpacity = 0.2, stroke = "none") |>
viz_graticule(step = 20, stroke = "#38896F",
              strokeDasharray = c(5,10)) |>
viz_render()
```

## Earth

The `viz_earth` function displays PNG files representing the Earth's surface. The data comes from Natural Earth. Credit as follows: Natural Earth. Free vector and raster map data @ naturalearthdata.com. Unlike tiles, you can choose whichever projection you want. the `url` parameter allows to change the image to display. You can choose one from this list: `"GRAY_50M_SR"`, `"GRAY_50M_SR_OB"`, `"GRAY_50M_SR_W"`, `"HYP_50M_SR"`, `"HYP_50M_SR_W"`, `"MSR_50M"`, `"NE1_50M_SR_W"`, `"NE2_50M_SR"`, `"NE2_50M_SR_W"`, `"OB_50M"`, `"PRIMSA_SR_50M"`, `"SR_50M"`. 

```{r}
viz_create(projection = "Polar") |>
viz_earth(url = "HYP_50M_SR_W") |>
viz_render()
```

## Tile

The `viz_tile` function allows function allows to add zoomable raster tiles. When you use this mark, the Mercator projection is automaticaly used. With the parmeter `url`, you can choose the style of the tiles. Try `"openstreetmap"`, `"opentopomap"`, `"worldterrain"`, `"worldimagery"`, `"worldStreet"`, `"worldphysical"`, `"shadedrelief"`, `"stamenterrain"`, `"cartodbvoyager"`, `"stamentoner"` , `"stamentonerbackground"` , `"stamentonerlite"`, `"stamenwatercolor"`, `"hillshade"`, `"worldocean"`, `"natgeo"` or `"worldterrain"`. 

```{r}
viz_create() |>
viz_tile(url = "worldimagery") |>
viz_render()
```

# Map Decorations and Annotations

## Labels

The `viz_text` function allows to add a text on the map. It allow also to create a layer with labels from a spatial dataframe. 

```{r}
africa <- world[world$region == "Africa",]
viz_create(projection = "Mercator", domain = africa) |>
viz_path(datum = africa, fill = "#CCC", fillOpacity = 0.5) |>
viz_text(data = africa, text = "name", fill = "#38896F",
         stroke = "white", strokeWidth = 4) |>
viz_render()
```

## Header

The `viz_header` function allows function allows to add a title above the map. You can customize the display using parameters such as `text`, `fill`, `fill_background`, `textAnchor`, `dominantBaseline` and many others.

```{r}
viz_create(projection = "EqualEarth") |>
viz_outline(fill = "#CCCCCC") |>
viz_path(data = world, fill = "white",
         fillOpacity = 0.2, stroke = "none") |>
viz_header(text = "Hello World", fill = "#38896F") |>
viz_render()
```

## Footer

The `viz_footer` function allows function allows to add a text below the map. You can customize the display using parameters such as `text`, `fill`, `fill_background`, `textAnchor`, `dominantBaseline` and many others.

```{r}
viz_create(projection = "EqualEarth") |>
viz_outline(fill = "#CCCCCC") |>
viz_path(data = world, fill = "white", fillOpacity = 0.2,
         stroke = "none") |>
viz_footer(text = "Author, source, date...",
           fill = "#38896F") |>
viz_render()
```

## North

The `viz_north` function adds a north arrow. By default, the arrow indicates true north at the location where it is placed. Depending on the map projection, the arrow may therefore not point in the same direction everywhere on the map. The `pos` parameter allows you to position the arrow using the SVG document coordinates (0,0 is at the top-left corner). 

```{r}
africa <- world[world$region == "Africa",]
viz_create(projection = "Mercator", domain = africa) |>
viz_path(datum = africa, fill = "#CCC", fillOpacity = 0.5) |>
viz_north(pos = c(50,50), fill = "#38896F") |>  
viz_render()
```

## Scalebar

The `viz_scalebar` function allows add a scalebar. The `pos` parameter allows you to position the scalebar.

```{r}
viz_create(projection = "Mercator", domain = africa) |>
viz_path(datum = africa, fill = "#CCC", fillOpacity = 0.5) |>
viz_scalebar() |>  
viz_render()
```

## Rhumbs

The `viz_rhumbs` function allows to display "rhumb lines" like on old portolan charts

```{r}
viz_create(projection = "Mercator", domain = africa) |>
viz_path(datum = africa, fill = "#CCC", fillOpacity = 0.5) |>
viz_rhumbs(pos = c(20,20), stroke = "#38896F", nb = 32) |>  
viz_render()
```

## Pattern

The `viz_pattern` function allows to display a pattern on the map

```{r}
viz_create(projection = "Equal Earth") |>
viz_outline(fill = "white") |>
viz_pattern(clipOutline = TRUE, stroke = "#38896F", pattern = "waves") |>
viz_path(datum = world, fill = "#CCC") |>
viz_render()
```

# Effects

The `geoviz` package makes it possible to create publication-quality maps. Since these maps are built in SVG format, SVG filters can be defined. This works in two steps. First, you define an effect by giving it an identifier. Then, you apply it to the layer of your choice. 


```{r}
aus <- world[world$ISO3 == "AUS",]
```

# blur

The `viz_blur` function is used to define a blur effect The `id` parameter defines the filter identifier. The `stdDeviation` parameter determines the blur range.

```{r}
# A simple map
viz_create() |>
  # 1 - create filter
  viz_blur(id = "my_blur_effect", stdDeviation =  5) |>
  # 2 - add path and apply filter
  viz_path(datum = aus, fill =  "#38896F",
           filter = "url(#my_blur_effect)") |>
  viz_render()
```

# shadow

The `viz_shadow` function can be used to draw a drop shadow

```{r}
viz_create(margin = 5) |>
  viz_shadow(id = "my_shadow_effect", stdDeviation =  2.5,
             dx = 5, dy = 5) |>
  viz_path(datum = aus, fill =  "#38896F",
           filter =  "url(#my_shadow_effect)") |>
  viz_render()
```

# radialGradient

With `viz_radialGradient`, you can create a radial gradient. In addition to the filter id, the function takes several parameters as input. `color1` and `color2` define the gradient colors. The parameters `offset1`, `offset2`, `fx` and `fy` define how the gradation is performed. Then, you have to apply it on the `fill` attribute.

```{r}
viz_create(margin = 5) |>
  viz_radialGradient(id = "my_radial_gradient",
  color1 = "#7db054", color2 = "#38896f",
  offset1 = 50, offset2 = 100, fx = 50, fy = 50) |>
  viz_path(datum = aus, fill = "url(#my_radial_gradient)") |>
  viz_render()
```

# clipPath

The <clipPath> SVG element defines a clipping path, to be used by the clip-path property. A clipping path restricts the region to which paint can be applied. In geoviz, you can add a clipPath element with the `viz_clipPath()` function.

```{r}
viz_create() |>
  viz_clipPath(id = "my_clip_path", datum =  aus )|>
  viz_tile(url = "worldimagery", clipPath = "url(#my_clip_path)") |>
  viz_render()
```

# pattern

The `viz_pattern` function (which we have already discussed) allows you to create patterns and apply them to any geometry. It supports multiple textures: `"lines"`, `"crosshatch"`, `"dots"`, `"waves"`, `"triangles"`, and `"zigzag"`.

```{r}
viz_create() |>
viz_pattern(data = aus, pattern = "cross",
            stroke = "#38896F", strokeOpacity = 2,
            angle = 45) |>
viz_render()
```

# sketch

The `viz_sketch` function generates a hand-drawn (sketchy) SVG representation of a spatial dataframe. It applies SVG filters (*feTurbulence* and *feDisplacementMap*) to create a pencil-like effect. 

```{r}
viz_create() |>
viz_sketch(data = aus, stroke = "#38896F") |>
viz_render()
```

# Design nice maps

With all these highly customizable layers, it is possible to create very nice maps with geoviz. Here is an example.

```{r}
viz_create(projection = "mercator", domain = africa, 
           margin = 20, zoomable = FALSE) |>
viz_clipPath(id = "africaclip", datum =  africa )|>
viz_shadow(id = "africashadow", stdDeviation =  4,
           dx = 4, dy = 4, opacity = 0.3) |>
viz_outline()|>
viz_path(datum = world, fill = "white", fillOpacity = 0.3) |>
viz_path(data = africa,   filter =  "url(#africashadow)") |>
viz_rhumbs(pos = c(0,-10), coords = "geo", stroke = "white",
           strokeWidth = 2, strokeDasharray = c(5,3)) |>
viz_path(data = africa, fill = "#d6bc47", fillOpacity = 1) |>
viz_tile(url = "hillshade", clipPath = "url(#africaclip)",
         opacity = 0.3) |>
viz_sketch(data = africa, simplify = 1, strokeOpacity = 0.6) |>
viz_header(text = "The African continent", fill = "white",
           background_fill = "black") |>
viz_pattern(stroke = "white", angle = 45) |>
viz_scalebar()|>
viz_north()|>
viz_render()
```
