---
title: "layout"
author: "Nicolas Lambert"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{layout}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, echo=FALSE, message=FALSE}
library(geoviz)
library(sf)
world <- st_read(system.file("gpkg/world.gpkg", package = "geoviz"), quiet = TRUE)
```

The geoviz library offers numerous functions that allow you to display specific layers and overlay them to create the map of your choice. We list them below.

# path

The `viz_path` function allows you to display any basemap (points, lines, and polygons). You can customize the display using SVG parameters such as `fill`, `stroke`, `strokeWidth`, `opacity`, `fillOpacity`, and [many others](https://riatelab.github.io/geoviz/global.html#path).

```{r}
viz_create(projection = "EqualEarth") |>
  viz_path(data = world, fill = "#38896F") |>
  viz_render()
```

# outline

The `viz_outline` function allows to create a layer with Earth outline in the projection. You can customize the display using SVG parameters such as `fill`, `stroke`, `strokeWidth`, `opacity`, `fillOpacity`, and [many others](https://riatelab.github.io/geoviz/global.html#outline).

```{r}
viz_create(projection = "EqualEarth") |>
  viz_outline(fill = "#38896F") |>
  viz_path(data = world, fill = "white", fillOpacity = 0.2, stroke = "none") |>
  viz_render()
```

# graticule

The `viz_graticule` function allows function allows to create a layer with lat/long lines. You can customize the display using SVG parameters such as `step`, `stroke`, `strokeWidth`, `strokeDasharray`, and [many others](https://riatelab.github.io/geoviz/global.html#graticule).

```{r}
viz_create(projection = "EqualEarth") |>
  viz_path(data = world, fill = "#38896F", fillOpacity = 0.2, stroke = "none") |>
  viz_graticule(step = 20, stroke = "#38896F", strokeDasharray = c(5,10)) |>
  viz_render()
```

# header

The `viz_header` function allows function allows to add a title above the map. You can customize the display using parameters such as `text`, `fill`, `fill_background`, `textAnchor`, `dominantBaseline` and [many others](https://riatelab.github.io/geoviz/global.html#header).

```{r}
viz_create(projection = "EqualEarth") |>
  viz_outline(fill = "#CCCCCC") |>
  viz_path(data = world, fill = "white", fillOpacity = 0.2, stroke = "none") |>
  viz_header(text = "Hello World", fill = "#38896F") |>
  viz_render()
```

# footer

The `viz_footer` function allows function allows to add a text below the map. You can customize the display using parameters such as `text`, `fill`, `fill_background`, `textAnchor`, `dominantBaseline` and [many others](https://riatelab.github.io/geoviz/global.html#footer).

```{r}
viz_create(projection = "EqualEarth") |>
  viz_outline(fill = "#CCCCCC") |>
  viz_path(data = world, fill = "white", fillOpacity = 0.2, stroke = "none") |>
  viz_footer(text = "Author, source, date...", fill = "#38896F") |>
  viz_render()
```

# tile

The `viz_tile` function allows function allows to add zoomable raster tiles. When you use this mark, the Mercator projection is automaticaly used. With the parmeter `url`, you can choose the style of the tiles. Try `"openstreetmap"`, `"opentopomap"`, `"worldterrain"`, `"worldimagery"`, `"worldStreet"`, `"worldphysical"`, `"shadedrelief"`, `"stamenterrain"`, `"cartodbvoyager"`, `"stamentoner"` , `"stamentonerbackground"` , `"stamentonerlite"`, `"stamenwatercolor"`, `"hillshade"`, `"worldocean"`, `"natgeo"` or `"worldterrain"`. See Other parameters [here](https://riatelab.github.io/geoviz/global.html#tile).

```{r}
viz_create() |>
  viz_tile(url = "worldimagery") |>
  viz_render()
```

# earth

The `viz_earth` function displays PNG files representing the Earth's surface. The data comes from Natural Earth. Credit as follows: Natural Earth. Free vector and raster map data @ naturalearthdata.com. Unlike tiles, you can choose whichever projection you want. the `url` parameter allows to change the image to display. You can choose one from this list: `"GRAY_50M_SR"`, `"GRAY_50M_SR_OB"`, `"GRAY_50M_SR_W"`, `"HYP_50M_SR"`, `"HYP_50M_SR_W"`, `"MSR_50M"`, `"NE1_50M_SR_W"`, `"NE2_50M_SR"`, `"NE2_50M_SR_W"`, `"OB_50M"`, `"PRIMSA_SR_50M"`, `"SR_50M"`. See Other parameters [here](https://riatelab.github.io/geoviz/global.html#earth).  

```{r}
viz_create(projection = "Polar") |>
  viz_earth(url = "HYP_50M_SR_W") |>
  viz_render()
```

# north


```{r}
africa <- world[world$region == "Africa",]
viz_create(projection = "Mercator", domain = africa) |>
  viz_path(data = africa, fill = "#38896F") |>
  viz_render()
```


# scale

```{r}
africa <- world[world$region == "Africa",]
viz_create(projection = "Mercator", domain = africa) |>
  viz_path(data = africa, fill = "#38896F") |>
  viz_render()
```

# hatch


# sketch

```{r}
viz_create(projection = "EqualEarth") |>
  viz_sketch(data = world, stroke = "#38896F") |>
  viz_render()
```
