---
title: "Projections"
author: "Nicolas Lambert"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Projections}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r}
#| include: false
knitr::opts_chunk$set(out.width = "100%")
```

# Preliminary remarks

The projections in geoviz are based on the ecosystem of D3.js. More specifically, they rely on the geographic projection functions implemented in D3, which are originally designed to be used in JavaScript.

These projection methods assume a spherical model of the Earth, rather than an ellipsoidal one. In other words, the Earth is treated as a perfect sphere, which simplifies the mathematical transformations required to project geographic coordinates (longitude and latitude) onto a 2D basemap.

Under the hood, three JavaScript libraries are used. Feel free to visit their respective pages to explore the available projections and learn their names.

- **d3.geo** https://d3js.org/d3-geo/projection
- **d3-geo-projection** https://github.com/d3/d3-geo-projection
- **d3-geo-polygon** https://github.com/d3/d3-geo-polygon

👉 *E.g. `d3.geoBertin1953()` in theses librairies become `"Bertin1953"` in `geoviz`!*

# In geoviz

In `geoviz`, the projection is defined in the `viz_create` function using the `projection` parameter. You simply need to specify a string corresponding to the desired projection.

You must therefore always provide a basemap in WGS84. The transformation is performed <ins>on the fly</ins> at display time.

For example:

```{r, echo=TRUE, message=FALSE}
library(geoviz)
library(sf)
world <- st_read(
  system.file("gpkg/world.gpkg", package = "geoviz"),
  quiet = TRUE
)
```

```{r}
viz_create(projection = "Bertin1953") |>
  viz_outline() |>
  viz_graticule(stroke = "white") |>
  viz_path(data = world, fill = "#38896F") |>
  viz_render()
```

There are many possibilities. For example, you can try "Polar", "EqualEarth", "Airocean", "PolyhedralWaterman", "InterruptedHomolosine", "Berghaus", etc.

```{r}
viz_create(projection = "PolyhedralWaterman") |>
  viz_outline() |>
  viz_graticule(stroke = "white") |>
  viz_path(data = world, fill = "#38896F") |>
  viz_render()
```

```{r}
viz_create(projection = "Spilhaus") |>
  viz_outline() |>
  viz_graticule(stroke = "white") |>
  viz_path(data = world, fill = "#38896F") |>
  viz_render()
```

# Custom projection

Each projection is a function that transforms the sphere onto a plane. However, it is always possible to configure it to change the projection center.

```{r}
viz_create(projection = "EqualEarth.rotate([0,90])") |>
  viz_outline() |>
  viz_graticule(stroke = "white") |>
  viz_path(data = world, fill = "#38896F") |>
  viz_render()
```

# Globe view

If you want to display your layers on a globe, simply choose an orthographic projection. If you want the globe to rotate, use `zoomable = "versor"` (named after the JavaScript library that enables this).

```{r}
viz_create(projection = "orthographic", zoomable = "versor") |>
  viz_outline() |>
  viz_graticule(stroke = "white") |>
  viz_path(data = world, fill = "#38896F") |>
  viz_render()
```

Note that `zoomable = "versor"` works alos wiyh other projections 🤩. Actually, it allows to change the center of the projection. 


```{r}
viz_create(projection = "EqualEarth", zoomable = "versor") |>
  viz_outline() |>
  viz_graticule(stroke = "white") |>
  viz_path(data = world, fill = "#38896F") |>
  viz_render()
```
